---
title: "Project memo"
author: "Team name"
output: github_document
always_allow_html: true
---


```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)
library(leaflet)
library(RColorBrewer)
library(htmlwidgets)

squirrels <- read_csv("../data/2018_Central_Park_Squirrel_Census_-_Squirrel_Data.csv")

```



## Loading the RColorBrewer pallette "Paired," which includes pairs of light
## and dark shades.

```{r color-brewer}

my_colors <- brewer.pal(n = 4, name = "Paired")


```



### Final Plot 1 - Leaflet Map:

To build the leaflet map, the first step was to create separate variables for
the latitude and longitude, as both were combined in the original dataset. This 
was the most intensive data cleaning necessary for the project. The code then 
generates the base OpenStreetMap layer, as well as grouped circle markers for 
each squirrel.

```{r leaflet-map}

# Cleaning the strange formatting from the original dataset
squirrels <- squirrels |>
  mutate(lng = substring(`Lat/Long`, 8, 22),
         lat = substring(`Lat/Long`, 26, 40))
squirrels$lng <- as.numeric(squirrels$lng)
squirrels$lat <- as.numeric(squirrels$lat)

# Resolving an issue with the substring function that sent 200 squirrels to 
# Colombia
squirrels <- squirrels |>
  mutate(
    lat = if_else(
      condition = lat < 10,
      true = lat + 40,
      false = lat
    )
  )

# Generating the leaflet map
squirrels_map <- leaflet(squirrels) |>
  addTiles() |>
  setView(lng = -73.9651,
          lat = 40.7833,
          zoom = 13) |>
  addCircleMarkers(lat = ~lat, 
             lng = ~lng, 
             clusterOptions = markerClusterOptions(),
             color = squirrels$`Primary Fur Color`,
             fillColor = squirrels$`Primary Fur Color`,
             popup = squirrels$`Unique Squirrel ID`)

# Saving the HTML widget
saveWidget(squirrels_map, file = "index.html", selfcontained = TRUE)
```

```{r squirrel-activities}
# Creating a "No Activity" variable
activities <- squirrels |>
  mutate("No Activity" = case_when (Running == FALSE & 
                                    Chasing == FALSE & 
                                    Climbing == FALSE & 
                                    Eating == FALSE & 
                                    Foraging == FALSE
                                    ~ TRUE,
                                  )) |>
  
# Counting the number of squirrels observed doing each activity
  select(Shift, Running, Chasing, Climbing, Eating, Foraging, "No Activity") |>
  pivot_longer(
    cols = c(Running, Chasing, Climbing, Eating, Foraging, "No Activity"),
    names_to = "Activity",
    values_to = "Observed"
  ) |>
  filter(Observed == TRUE) |>
  count(Activity, Shift, name = "Frequency") |>
  group_by(Activity) |>
  mutate(Total_Frequency = sum(Frequency)) |>
  ungroup() |>
  
# Expressing with percentages
  group_by(Shift) |>
  mutate(Percent = (100*(Frequency / sum(Frequency)))) |>
  ungroup()

# Generating the plot with paired bars dividing the AM and PM shifts
ggplot(activities, aes(x = reorder(Activity, -Total_Frequency), 
                      y = Percent, fill = Shift)) +
  geom_col(position = "dodge") +
  labs(
    title = "Comparison of Squirrel Activities",
    x = "Squirrel Behavior",
    y = "Percentage"
  ) +
  
# Using a more distinctive color palette
  scale_fill_brewer(palette = "Paired")

# Saving the plot
  ggsave("squirrelActivity.png", plot = get_last_plot(), width = 6, height = 3)
```


#### Final Plot 3

```{r squirrel-human-behavior}

# A similar bar plot to investigate squirrel-human interactions
squirrels %>%
  select(Shift, Approaches, Indifferent, `Runs from`) %>%
  pivot_longer(
    cols = c(Approaches, Indifferent, `Runs from`),
    names_to = "Interaction",
    values_to = "Observed"
  ) %>%
  filter(Observed == TRUE) %>%
  count(Shift, Interaction, name = "Frequency") %>%
  
# Constructing plot divided by AM and PM shift
  ggplot(aes(
    x = Interaction,
    y = Frequency,
    fill = Shift 
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Squirrel-Human interactions by time of shift (AM or PM)",
    x = "Squirrel Behavior",
    y = "Number of Squirrels Observed Exhibiting Behavior",
    fill = "Time of shift"
  ) 

ggsave("squirrelHumanActivity.png", plot = get_last_plot(), width = 6, height = 3)

```


